---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

**Setting global option for echo = TRUE for the code to appear**
```{r setoptions , echo = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE , message=FALSE )

```

## 1.Unzipping , Loading and preprocessing the data

```{r}
if (!file.exists("activity.csv")){
  unzip(zipfile = "activity.zip")
}
```

### Reading the data and explorting a little

```{r}
activity = read.csv("activity.csv")

head(activity)
summary(activity)

```

## 2.Histogram of the total steps taken per day
```{r}

library(dplyr)
steps_per_day = activity[,c(1,2)] %>% group_by(date) %>% summarise(steps = sum(steps , na.rm = T))
head(steps_per_day)

library(ggplot2)
ggplot(steps_per_day ,aes(steps))+
geom_histogram(color="darkblue", fill="lightblue")+
ggtitle("Histogram of the total steps taken per day")+
ylab("Frequency")
```  
<br>


## 3.Mean and median number of steps taken each day
```{r}
steps_per_day_mean_median = activity[,c(1,2)] %>% group_by(date) %>% summarise(mean = mean(steps , na.rm = T) , median = median(steps , na.rm = T))
head(steps_per_day_mean_median)

mean(steps_per_day$steps , na.rm = T)
median(steps_per_day$steps , na.rm = T)

```

## 4.Time series plot of the average number of steps taken
```{r}
intervals_average_steps = activity[,c(1,3)] %>% group_by(interval) %>% summarise(mean = mean(steps , na.rm = T) )
plot (intervals_average_steps$interval , intervals_average_steps$mean,
      type = "l",
      main = "The average daily activity pattern",
      xlab = "5 min intervals",
      ylab = "average number of steps taken")

```  
<br>



## 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_interval = intervals_average_steps$interval[which.max(intervals_average_steps$mean)]
max_interval
```

## 6.Code to describe and show a strategy for imputing missing data

#### **Number of missing values and percentage to whole data**
```{r}
sum(is.na(activity))
paste(round(sum(is.na(activity))/nrow(activity) * 100 ,2 ) ,"%" ,sep = "")

```

#### **Impute missing values in data with the average steps for each day**
```{r}
new_activity = activity
activity_NA = activity[is.na(activity),]
activity_with_mean = merge (activity_NA,intervals_average_steps)
new_activity$steps[is.na(activity$steps)] = activity_with_mean$mean 

```

#### **Testing the new data**

```{r}
sum(is.na(new_activity)) #Empty !

```

## 7.Histogram of the total number of steps taken each day after missing values are imputed

```{r}
new_steps_per_day = new_activity[,c(1,2)] %>% group_by(date) %>% summarise(steps = sum(steps))
head(new_steps_per_day)

ggplot(new_steps_per_day ,aes(steps))+
  geom_histogram(color="darkblue", fill="lightblue")+
  ggtitle("Histogram of the total steps taken per day after removing NAs")+
  ylab("Frequency")

```

#### **Histogram of both the original data and the data after removing NAs to compare between both**

```{r}
new_steps_per_day$type = "new"
steps_per_day$type = "original"
both  = rbind(new_steps_per_day, steps_per_day)

ggplot(both, aes(steps, fill = type)) + 
  geom_histogram(alpha = 0.5, position = 'identity')

```  

<br>

#### **Summary of the original and imputed data**

```{r}
summary(new_steps_per_day$steps) #Original
summary(steps_per_day$steps)[-7] #imputed
```
We can see that the main difference is in the fisrt quantile  


## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
new_activity$date = as.Date(new_activity$date)
new_activity$daytype = ifelse(weekdays(new_activity$date)=="Saturday" | weekdays(new_activity$date)=="Sunday", "Weekend", "Weekday")


intervals_average_steps_2groups = new_activity[,c(1,3,4)] %>% group_by(interval,daytype) %>% summarise(mean = mean(steps))

ggplot(intervals_average_steps_2groups, aes(intervals_average_steps_2groups$interval , intervals_average_steps_2groups$mean ,col = daytype ))+
  geom_line()+
  facet_wrap(daytype ~ . , ncol = 1, nrow=2)+
  xlab("5 min intervals")+ylab("average number of steps taken")+
  ggtitle ("The average daily activity pattern")+
  scale_colour_brewer(palette="Set1" )

```

#### **9.All the code can bee see as 'Echo' is set to TRUE**
